### 2021-04-16 manuscript.rmd
Replace all 'comparative content analysis'

---
title             : "The Bayesian turn in comparative content analysis"
shorttitle        : "BAYESIAN MULTILEVEL REGRESSION"

@@

abstract: |
  Comparative content analysis is a frequently used research design in journalism studies. The purpose of this paper is to argue that researchers should refrain from using current (frequentist) analytical approaches to analyze data from a comparative content analysis. It is because current approaches cannot account for non-atomicity (nested nature) and non-stochasticity (non-random sampling) of the data. Bayesian multilevel regression is suggested to be a, and probably, the only, valid analytical approach. Using the openly available useNews data and the R package brms, we demonstrate how to apply such approach for the analysis of comparative content analysis. We also address common queries of the approach such as choosing the prior distribution and diagnosing the convergence of the model. Furthermore, we argue that choosing a Bayesian approach also solves the problem of often misinterpreted frequentist confidence intervals.

keywords          : "Bayesian inference, multilevel model, comparative content analysis, ecological effect"
wordcount         : "0"

@@

Comparative content analysis is a frequently used research design in journalism studies. Comparative content analysis is used in modeling journalistic culture [e.g. @esser:2013:C], comparative news value research [e.g. @burggraaff:2017:T;@wilke:2012], news flow research [e.g. @wu:2000:SDI;@grasland:2019:I], among others. A typical comparative content analysis in journalism research includes news articles from various outlets and usually also from various countries. Then the included news articles are coded either manually or automatically. Although some comparative content analysis studies are purely descriptive, most of these studies are aimed to study how contextual conditions such as characteristics of media outlets (e.g. political orientation, online versus offline) and their countries (e.g. post-communist country) influence the content of news articles. For example, some hypotheses are "the coverage of foreign countries by the news is primarily determined by geographical proximity and the status of the covered country" [@wilke:2012, p.306] and "the  degree of opinion-orientation will be highest in newspapers from Polarized Mediterranean systems and lowest in those from Anglo-American systems." [@esser:2013:C] Employing the language from epidemiology [@susser:1994], the effect on media content in these hypotheses is assumed to be ecological. It assumes a macro contextual factor (e.g. newspapers from Polarized Mediterranean systems) is associated with an outcome at the micro-level (e.g. degree of opinion-orientation), namely, the article-level.

This study of ecological effect on individual behaviors has a long tradition in social sciences. A famous example was Émile Durkheim's investigation on the ecological factors associated with suicide [@durkheim1897suicide]. Durkheim found the higher suicide rate among citizens from Scandinavian countries than those from other European countries. Using today’s standard, Durkheim's analysis is descriptive and this descriptive approach for studying ecological effect is indeed still useful for hypothesis generation in comparative content analysis. An example of hypothesis-generating comparative content analysis is @chan2020combining, in which the sentiment profiles of terrorism coverage from Muslim- and Christian-majority countries were visualized. Hypothesis-generating comparative content analysis, however, is rare. Most studies have a confirmatory outlook. As we see from the example hypotheses above, most of these studies propose hypotheses to test for ecological effects. In this context usually traditional (frequentist) hypothesis testing approaches were used: @esser:2013:C lumped multiple outlets from the same country together and then used univariate analysis of variance to test for the differences in the proportion of opinion-orienting articles across countries. In @wu:2000:SDI, multiple stepwise regression was used. These approaches violate the underlying assumptions of the statistical analyses. It highlights the fact that comparative content analysis introduces a feature that researchers usually overlook: media contents are clustered in multilevel. A news article is nested within its media outlet and its media outlet is in turn nested within its country. Theoretically with a large enough sample it can be argued that a news article is nested within authors. Such data structure brings two problems: non-atomicity and non-stochasticity.

In our paper we first discuss these two issues before we suggest Bayesian multilevel regression as a standard approach for comparative research in journalism studies. While Bayesian models might be slightly more complicated to estimate in contrast to frequentist models as priors have to be chosen, these models allow researchers to make "principled statements about precision and plausibility" [@morey:2015 p. 121] and solve the issue of misinterpreted confidence intervals [@hoekstra:2014:R;@rinke:2018:PMA]. Using the openly available useNews data and the R package brms, we then demonstrate how to apply a Bayesian approach for the analysis of comparative content analysis.

@@

In a comparative content analysis, it is easy to incorrectly assume that macro-level independent variables (e.g. Anglo-American systems) could be analyzed at the same level as the micro-level dependent variable (e.g. the  degree of opinion-orientation). The manifestation of this incorrect assumption is to enter macro-level variables as independent variables in multiple regression analysis and regress them against a micro-level dependent variable. Suppose one wants to study how the probability of covering China changes when the hosting country of the media outlet has a higher trade volume with China. In the following regression equation, let $y_{i}$ be the dependent variable at the micro-level (coverage of China or not) and $x_{i}$ be the independent variable at the macro-level (trade volume with China). Suppose there are $m$ news articles where $i = 1,2,\ldots,m$.

@@

Multilevel modeling is the way to go for analyzing non-atomic data from comparative content analysis. However, the conventional approach for statistical inference makes it not optimal.

@@

Comparative content analysis, unlike randomized surveys or randomized media experiments, usually collects all available data. In contrast to sampling, these studies are actually doing census of all data and there is no way to get more data unless the scope of these studies is changed [@berk:1995:SIA]. It is especially true for modern large-N studies using automated content analytic techniques. @burggraaff:2017:T, for instance, "collected all available news items from a selection of major Dutch news outlets, both online and print" (p. 6) and that amounted to 762,095 articles from 9 outlets in the period of 2014-2015. 
Census-style studies are not repeatable and thus they generate non-stochastic data. In other words, data from comparative content analysis are fundamentally irrelevant for frequentist inference [@western:1994:BIC]. P-values and confidence intervals generated do not have the same meaning as those from repeatable studies. According to @western:1994:BIC [p. 413], these values from non-stochastic data "lack meaning even as abstract propositions".

Following arguments from @western:1994:BIC and @stegmueller:2013:HMC for comparative studies, Bayesian inference should be used for analyzing data from comparative content analysis. Choosing a Bayesian approach also solves the problem of misinterpreted confidence intervals [@hoekstra:2014:R;@rinke:2018:PMA]. Switching to a Bayesian approach addresses this issue as Bayesian "credible intervals support an interpretation of probability in terms of plausibility" [@morey:2015, p.120]. However, for Bayesian models priors have to be chosen by researchers.

@@

Bayesian analysis is still a minority statistical method in social sciences. It is reasonable to say that Bayesian analysis is extremely rare in our field. Although some communication researchers adopted the method [e.g. @kovic:2017:B;@chan2020high] these studies are not comparative. As far as we know, the only available comparative content analysis that Bayesian multilevel regression was used for studying ecological effect is @leeuw:2020:AAT. We believe comparative content analysis fits the use case of Bayesian multilevel regression analysis. Although there is another methodological introduction to Bayesian analysis aiming at communication researchers [@konijn:2015:PSP], the main purpose of this paper is to demonstrate how to practically implement one using the R package *brms* [@burkner2017advanced]. As the interface of *brms* is almost the same as *lme4* [@bates:2015:FLM, another R package for fitting multilevel models using MLE], *lme4* users might find *brms* extremely familiar. The above quoted challenge about the prescription of more widespread use of Bayesian methods for multilevel modeling is no longer an issue.

@@

In this paper, we argue the case for using Bayesian multilevel regression analysis to analyze data from comparative content analyses. We demonstrate using the openly available useNews dataset that Bayesian analysis provides valid inference of ecological effects and can be done easily with the R package *brms*. Using a Bayesian approach, however, has also clearly some disadvantages. Researchers have to define priors in order to run a Bayesian regression analysis. As we could show this limitation does not pose a controversial challenge as weakly informative priors can be chosen and in most cases enough data is available and thus the priors have almost no influence on the posterior distribution of the parameters. If we have prior knowledge it allows us to create even better models. Furthermore, it takes more time to get the results of a Bayesian regression analysis. Estimating a Bayesian regression model is a computationally demanding task. It took 4 days on a regular computer to get the results of the Bayesian regression model with 1,525,871 articles (Model B1). Still, eventually it worked and typical studies in comparative journalism research [e.g. 6525 articles in @esser:2013:C] can be estimated in a reasonable time frame [^TIME].

>>>>

---
title             : "The Bayesian turn in comparative communication research"
shorttitle        : "BAYESIAN MULTILEVEL REGRESSION"

@@

abstract: |
  Comparative approach such as comparative content analysis is frequently used in journalism studies. The purpose of this paper is to argue that researchers should refrain from using current (frequentist) analytical approaches to analyze data from a comparative study. It is because current approaches cannot account for non-atomicity (nested nature) and non-stochasticity (non-random sampling) of the data. Bayesian multilevel regression is suggested to be a ---and probably, the only--- valid analytical approach. Using the openly available *useNews* data and the R package *brms*, we demonstrate how to apply such approach for the analysis of comparative data. We also address common queries of the approach such as choosing the priors and diagnosing the convergence of the model. Furthermore, we argue that choosing a Bayesian approach also solves the problem of often misinterpreted frequentist confidence intervals.

keywords          : "Bayesian inference, multilevel model, comparative communication research, ecological effect"
wordcount         : "0"

@@

Communication research is often comparative, journalism studies are no exception. For example, comparative approach is used in modeling journalistic culture [e.g. @esser:2013:C], comparative news value research [e.g. @burggraaff:2017:T;@wilke:2012], news flow research [e.g. @wu:2000:SDI;@grasland:2019:I], among others. A typical comparative study in journalism research, otherwise known as comparative content analysis, includes news articles from various outlets and usually also from various countries. Then the included news articles are coded either manually or automatically. Although some comparative studies are purely descriptive, most of these studies are aimed to study how contextual conditions such as characteristics of media outlets (e.g. political orientation, online versus offline) and their countries (e.g. post-communist country) influence the content of news articles. For example, some hypotheses are "the coverage of foreign countries by the news is primarily determined by geographical proximity and the status of the covered country" [@wilke:2012, p.306] and "the  degree of opinion-orientation will be highest in newspapers from Polarized Mediterranean systems and lowest in those from Anglo-American systems." [@esser:2013:C] Employing the language from epidemiology [@susser:1994], the effect on media content in these hypotheses is assumed to be ecological. It assumes a macro contextual factor (e.g. newspapers from Polarized Mediterranean systems) is associated with an outcome at the micro-level (e.g. degree of opinion-orientation), namely, the article-level.

This study of ecological effect on individual behaviors has a long tradition in social sciences. A famous example was Émile Durkheim's investigation on the ecological factors associated with suicide [@durkheim1897suicide]. Durkheim found the higher suicide rate among citizens from Scandinavian countries than those from other European countries. Using today’s standard, Durkheim's analysis is descriptive and this descriptive approach for studying ecological effect is indeed still useful for hypothesis generation in comparative research. An example of hypothesis-generating comparative research is @chan2020combining, in which the sentiment profiles of terrorism coverage from Muslim- and Christian-majority countries were visualized. Hypothesis-generating comparative study, however, is rare. Most studies have a confirmatory outlook. As we see from the example hypotheses above, most of these studies propose hypotheses to test for ecological effects. In this context usually traditional (frequentist) hypothesis testing approaches were used: @esser:2013:C lumped multiple outlets from the same country together and then used univariate analysis of variance to test for the differences in the proportion of opinion-orienting articles across countries. In @wu:2000:SDI, multiple stepwise regression was used. These approaches violate the underlying assumptions of the statistical analyses. It highlights the fact that comparative research introduces a feature that researchers usually overlook: media contents are clustered in multilevel. A news article is nested within its media outlet and its media outlet is in turn nested within its country. Theoretically with a large enough sample it can be argued that a news article is nested within authors. Such data structure brings two problems: non-atomicity and non-stochasticity.

In our paper we first discuss these two issues before we suggest Bayesian multilevel regression as a standard approach for comparative research in journalism studies. While Bayesian models might be slightly more complicated to estimate in contrast to frequentist models as priors have to be chosen, these models allow researchers to make "principled statements about precision and plausibility" [@morey:2015 p. 121] and solve the issue of misinterpreted confidence intervals [@hoekstra:2014:R;@rinke:2018:PMA]. Using the openly available *useNews* data and the R package *brms*, we then demonstrate how to apply a Bayesian approach for the analysis.

@@

In comparative research, it is easy to incorrectly assume that macro-level independent variables (e.g. Anglo-American systems) could be analyzed at the same level as the micro-level dependent variable (e.g. the  degree of opinion-orientation). The manifestation of this incorrect assumption is to enter macro-level variables as independent variables in multiple regression analysis and regress them against a micro-level dependent variable. Suppose one wants to study how the probability of covering China changes when the hosting country of the media outlet has a higher trade volume with China. In the following regression equation, let $y_{i}$ be the dependent variable at the micro-level (coverage of China or not) and $x_{i}$ be the independent variable at the macro-level (trade volume with China). Suppose there are $m$ news articles where $i = 1,2,\ldots,m$.

@@

Multilevel modeling is the way to go for analyzing non-atomic data from comparative research. However, the conventional approach for statistical inference makes it not optimal.

@@

Comparative research in journalism studies, unlike randomized surveys or randomized media experiments, usually collects all available data. In contrast to sampling, these studies are actually doing census of all data and there is no way to get more data unless the scope of these studies is changed [@berk:1995:SIA]. It is especially true for modern large-N studies using automated content analytic techniques. @burggraaff:2017:T, for instance, "collected all available news items from a selection of major Dutch news outlets, both online and print" (p. 6) and that amounted to 762,095 articles from 9 outlets in the period of 2014-2015.

Census-style studies are not repeatable and thus they generate non-stochastic data. In other words, data from comparative journalism studies are fundamentally irrelevant for frequentist inference [@western:1994:BIC]. P-values and confidence intervals generated do not have the same meaning as those from repeatable studies. According to @western:1994:BIC [p. 413], these values from non-stochastic data "lack meaning even as abstract propositions".

Following arguments from @western:1994:BIC and @stegmueller:2013:HMC for comparative studies, Bayesian inference should be used for analyzing data from comparative research in our field. Choosing a Bayesian approach also solves the problem of misinterpreted confidence intervals [@hoekstra:2014:R;@rinke:2018:PMA]. Switching to a Bayesian approach addresses this issue as Bayesian "credible intervals support an interpretation of probability in terms of plausibility" [@morey:2015, p.120]. However, for Bayesian models priors have to be chosen by researchers.

@@

Bayesian analysis is still a minority statistical method in social sciences. It is reasonable to say that Bayesian analysis is extremely rare in our field. Although some communication researchers adopted the method [e.g. @kovic:2017:B;@chan2020high] these studies are not comparative. As far as we know, the only available comparative research in journalism studies that Bayesian multilevel regression was used for studying ecological effect is @leeuw:2020:AAT. We believe comparative communication research fits the use case of Bayesian multilevel regression analysis. Although there is another methodological introduction to Bayesian analysis aiming at communication researchers [@konijn:2015:PSP], the main purpose of this paper is to demonstrate how to practically implement one using the R package *brms* [@burkner2017advanced]. As the interface of *brms* is almost the same as *lme4* [@bates:2015:FLM, another R package for fitting multilevel models using MLE], *lme4* users might find *brms* extremely familiar. The above quoted challenge about the prescription of more widespread use of Bayesian methods for multilevel modeling is no longer an issue.

@@

In this paper, we argue the case for using Bayesian multilevel regression analysis to analyze data from comparative communication research. We demonstrate using the openly available useNews dataset that Bayesian analysis provides valid inference of ecological effects and can be done easily with the R package *brms*. Using a Bayesian approach, however, has also clearly some disadvantages. Researchers have to define priors in order to run a Bayesian regression analysis. As we could show this limitation does not pose a controversial challenge as weakly informative priors can be chosen and in most cases enough data is available and thus the priors have almost no influence on the posterior distribution of the parameters. If we have prior knowledge it allows us to create even better models. Furthermore, it takes more time to get the results of a Bayesian regression analysis. Estimating a Bayesian regression model is a computationally demanding task. It took 4 days on a regular computer to get the results of the Bayesian regression model with 1,525,871 articles (Model B1). Still, eventually it worked and typical studies in comparative journalism research [e.g. 6525 articles in @esser:2013:C] can be estimated in a reasonable time frame [^TIME].
